Proposed Solution

### 2017-01-20 CAV17/abs.tex

\app is capable of detecting various errors that cannot
be detected by previous efforts,
including entry ordering errors, fine-grained value correlation errors, 
and entry missing errors. 
We evaluate \app using a real-world dataset with 261 incorrect 
MySQL configuration files,
and \app is able to correctly 
detect errors (previous work failed to
detect) in 217 files.

>>>>

\app is capable of detecting various errors that cannot
be detected by previous efforts,
including entry ordering errors, fine-grained value correlation errors, 
and missing entry errors. 
We evaluate \app using a real-world dataset with 261 incorrect 
MySQL configuration files,
and \app is able to correctly 
detect errors (previous work failed to
detect) in 217 files.

### 2017-01-20 CAV17/intro.tex

Configuration errors are one of the most important root causes of today's software system failures~\cite{xu15systems, yin11anempirical}.
In their empirical study, Yin {\em et al.}~\cite{yin11anempirical} report that about 31\% of system failures were caused by misconfiguration problems and only 20\% were caused by bugs in program code. 
Misconfigurations, result in various system-wide problems, such as security vulnerabilities, application crashes, severe disruptions in software functionality, and incorrect program executions~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.

While many efforts have been proposed to check, troubleshoot, diagnose, and repair configuration errors~\cite{attariyan10automating,su07autobash, whitaker04configuration},
  those tools mainly post-factum and cannot give the safety guarantees available from powerful static analysis tools in program verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15}.
Existing techniques for program verification do not fit this application domain, as configuration files often lack specifications and structure.
There are rarely formal, or any, specifications on the constraints of the key values.
In the case of MySQL, the canonical specification is a 5240 page reference manual~\cite{mysql-manual15} to describe valid configurations and behavior.
We aim to automatically learn a formal specification the encodes this information by examining a set of example configuration files.

In previous work~\cite{santolucitoCAV}, we used a set of manually verified files by an domain expert order to learn specifications.
This approach is expensive and still prone to error by the expert.
Related work on configuration verification has used machine learning

While the grammar of configuration files tends to be simple (a series of key value pairs with some additional context), the keys and values themselves lack often structure and are untyped plain text.

   1) lack of a specification which would describe properties of configuration files, and 
   2) a program structure of configuration files -- they are mainly a sequence of entries assigning some value to system variables. 
The language in which configuration files are written does not adhere to a specific grammar or syntax. 
In particular, the entries in configuration files are untyped. 
Moreover, there are surprisingly few rules specifying constraints on entries and there is no explicit structure policy for the entries.
Thus, automated verification of configuration files would be highly desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

@@

In this paper, we present \app, the first automatic verification framework
for general software configurations. \app is based on a collection of powerful learning 
algorithms that do not necessarily depend on templates. The learning process takes as input a
 large sample of configuration files. The process is language-agnostic and works for any kind of configuration file, 
but all of the files in the sample need to be of the same kind (such as
MySQL or HTTPD configuration files).
From that sample,
\app learns an abundant set of rules specifying various properties that hold on the given sample. The files in the sample might contain
errors -- we are, therefore, using probabilistic learning to derive a set of accurate rules.

@@

Our main contributions are as follows:

\item Using a probabilistic logical inference approach to the learning task, which builds a more informative structure than state of the art machine learning algorithms

\item A basic application of graph analysis on the learned model to improve results in a way not available to traditional machine learning approaches

\item An open-source implementation, \app, of this approach applied to a set of 256 real world benchmarks. \app identify XX new misconfigurations errors on YY of these files.
\end{enumerate}

>>>>

Configuration errors are one of the most important root causes of
today's software system failures~\cite{xu15systems, yin11anempirical}.
In their empirical study Yin {\em et al.}~\cite{yin11anempirical} report
that about 31\% of system failures were caused by misconfiguration problems and only 20\% were caused by bugs in program code. 
Misconfigurations, in practice, may result in various
system-wide problems, such as security vulnerabilities, 
application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.

While many efforts have been proposed 
to check, troubleshoot, diagnose, and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration},
those tools mainly try to understand {\emph{what}} caused the 
error -- they are still not on a level of
automatic verification tools used for regular program 
verification~\cite{Leino10Dafny, PiskacWZ14, BobotFMP15} that can
detect errors without executing the code.
Two main obstacles why we cannot simply apply the existing automatic 
tools and techniques to verification of configuration files are: 1) a lack
of a specification which would describe properties of configuration files, and 2) a program structure of configuration files -- they
are mainly a sequence of entries assigning some value to system variables. The language in which configuration files are written does 
not adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. Moreover, there are surprisingly few rules specifying constraints on entries and there
is no explicit structure policy for the entries.
Thus, automated verification of configuration 
files would be highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

@@

In this paper, we present \app, the first automatic verification framework
for general software configurations. 
\app is based on a collection of powerful learning 
algorithms that do not necessarily depend on explicit templates. 
The learning process takes as input a
 large sample of configuration files. The process is language-agnostic and works for any kind of configuration file, 
but all of the files in the sample need to be of the same kind (such as
MySQL or HTTPD configuration files).
From that sample,
\app learns an abundant set of rules specifying various properties that hold on the given sample. The files in the sample might contain
errors -- we are, therefore, using probabilistic learning to derive a set of accurate rules.

@@

### 2017-01-20 CAV17/motiv.tex

In this section we illustrate functionality of \app on several non-trivial configuration errors
extracted from real-world examples. Although the errors are relatively simple, we call them 
non-trivial, because the majority of existing tools, \eg, learning-based checking
tools~\cite{zhang14encore, wang04automatic}, cannot detect
these configuration errors. Most of the presented examples were found on StackOverflow,
a popular question and answer website for programmers. To better understand problems that users have with 
configuration files, we explored and analyzed misconfigurations on a large number of user forums and on-line discussion sites.

\para{Example~1: Ordering error.} 
Ordering errors were reported by Yin {\em et al.}~\cite{yin11anempirical} and our first
example illustrates how ordering errors can cause a system to crash. When a user configures PHP 
to run with the
Apache HTTP Server, most likely the user will take some already existing configuration files and adapt them
to suit her needs. The configuration file might contain, among others, 
the following lines:
\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
extension = mysql.so
...
extension = recode.so
\end{lstlisting}

@@

By invoking \app, the user can detect such a configuration error.
In particular, \app reports that {\tt recode.so} 
should appear before {\tt mysql.so}, as shown
below:

@@

\para{Example~2: Fine-grained value correlation error.} 
Our next real-world misconfiguration example~\cite{correlation} comes from a discussion on StackOverflow.
The user has configured her MySQL as in the following:

@@

Our tool can learn this simple correlation as well, but learning more complex properties requires a different 
approach to the learning 
process. We need to track several variables. It is not enough just to compare them, but we need to learn, as in this particular case, which
invariant must be preserved. It was pointed out by Xu {\em et al.}~\cite{xu15hey} that
detecting fine-grained value correlation errors 
is a much more challenging
task than the normal value correlation problem.

To the best of our knowledge, \app is the first tool that
is able to check such fine-grained value correlation problems.

\para{Example~3: Missing entry error.} 
Many critical system outages result from the fact that an important
entry was missing in the configuration file. 
We call such a problem a {\em missing entry error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world missing entry error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

\para{Example~4: Singular value anomalies.} 
Many system performance problems are caused by the
anomaly that a value is set either too high or too low.
For example, a parameter relating to memory might be too large,
exhausting the RAM and causing extreme slowness or even a crash. 
Consider the following real-world misconfiguration file%
~\cite{singleValue}: a user was experiencing the ``My SQL Server 
has gone away" error.  
This is difficult to debug, since the error message is not specific.
It turns out the following line in her configuration file was problematic:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
max_allowed_packet = 100M
\end{lstlisting}

The user eventually resolved the issue by replacing the above line with {\tt max\_allowed\_packet=2M}.
One possible way to reach this conclusion more quickly is to determine that the original value, 100 MB, 
is statistically deviant and extremely large for the {\tt max\_allowed\_packet} attribute. 
ConfigV can detect these
outliers. Run on the user's original configuration file, ConfigV would output:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
WARNING: Violated Upper Hampel Rule for max_allowed_packet with value 104857600 
\end{lstlisting}

In the above, 104857600 represents the integer value of 100 megabytes. The idea is to output
a warning to the user that her value falls outside of a range considered normal, so that she
can make the appropriate adjustment to the value herself. In the above situation, it is 
prudent to set the value of {\tt max\_allowed\_packet} lower. We call such a value a 
\emph{singular value anomaly}.

\para{Other errors.}
In addition to the given motivating examples \app can also detect and report configuration errors that can be exposed by
existing work, such as EnCore~\cite{zhang14encore} and
CODE~\cite{yuan11context} (a full scope of \app's capabilities is given in $\S$\ref{sec-eval}).
A recently reported configuration error made the MySQL
daemon fail to start~\cite{syserror}.
One entry was written as 
{\tt datadir=/root/appfinder/mysql}, and the type, as well as 
format, seemed correct.
However, the problem is that this directory should not 
contain the root directory; the correct entry
should be {\tt datadir=/appfinder/mysql}.
Such an error is called a system environment-related configuration error.
Both \app and EnCore can handle misconfigurations concerning system execution.
Furthermore, \app is also able to detect type errors and syntax errors
in configuration files. An example of a type error is assigning to a variable {\tt {general\_log}} 
a file path {\tt {/var/log/mysql/mysql.log}}. \app successfully reports this typing error: 
\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
TYPE ERROR: Expected a Int with P=1.0 for "general_log[mysqld]"
\end{lstlisting}

>>>>

In this section we present the capability of \app through 
detecting errors in several real-world misconfiguration examples. 
These examples were non-trivial configuration errors
that were reported on StackOverflow~\cite{stackoverflow},
a popular question and answer website for programmers and administrators. 
%To better understand these problems, 
%we explored and analyzed misconfigurations 
%on a large number of user forums and on-line discussion sites.

\para{Example~1: Ordering error.} 
Ordering errors were reported by Yin {\em et al.}~\cite{yin11anempirical} and our first
example illustrates how ordering errors can cause a system to crash. When a user configures PHP 
to run with the
Apache HTTP Server, most likely the user will take some already existing configuration files and adapt them
to suit her needs. The configuration file might contain, among others, 
the following lines:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
    extension = mysql.so
        ...
    extension = recode.so
\end{lstlisting}

@@

By invoking \app, the user can detect such a configuration error.
In particular, \app reports that {\tt recode.so} 
should appear before {\tt mysql.so}, as shown
below: \ennan{Mark, please update all the motivating examples'
output results according
to our newest version implementation.}

@@

\para{Example~2: Fine-grained value correlation error.} 
Our second misconfiguration example~\cite{correlation} 
comes from a discussion on StackOverflow.
The user has configured her MySQL as in the following:

@@

Our tool can learn this simple correlation as well, but learning more complex properties requires a different 
approach to the learning 
process. We need to track several variables. It is not enough just to compare them, but we need to learn, as in this particular case, which
invariant must be preserved. It was pointed out by Xu {\em et al.}~\cite{xu15hey} that
detecting fine-grained value correlation errors 
is a much more challenging
task than the normal value correlation problem.
To the best of our knowledge, \app is the first tool that
is able to check such fine-grained value correlation problems.

\para{Example~3: Missing entry error.} 
Many critical system outages result from the fact that an important
entry was missing from the configuration file. 
We call such a problem a {\em missing entry error}.
In a public misconfiguration dataset~\cite{configdataset},
many MySQL failure reports were caused by
missing entry errors.
Below is a real-world missing entry error example~\cite{yin11anempirical}:
when a user wants to use OpenLDAP to enable her directory access
protocol, she needs to use the password policy overlay. This is usually
achieved via the following entries in the OpenLDAP configuration file:

@@

\para{Example~4: Type errors.} 
Many system performance problems are caused by 
assigning incorrect type of values to some key in configuration
files. Consider the following real-world misconfiguration file: 
a user tries to install MySQL and she needs to initiate the path
of the log information generated by MySQL.
This user puts the following assignment in her MySQL
configuration file:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
    general_log = /var/log/mysql/mysql.log
\end{lstlisting}

Unbeknowest to this user, the entry ``general\_log'' should be an 
integer, not a string. In MySQL, there is another entry named
``general\_log\_file'' used to specify the log path.
With \app, this user can get the following result:

\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
    TYPE ERROR: Expected a Int with P=1.0 for
    "general_log [mysql]"
\end{lstlisting}

%The above result means that we need to assign an integer value to
%the entry ``general\_log''.
